---
title: "454_Genetic_Algorithms"
author: "Michael Gilbert"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    fig_caption: yes
    fig_crop: no
    fig_height: 4.75
    fig_width: 5.75
    highlight: tango
    number_sections: yes
  pdf_document: default
geometry: margin = 0.5in
---
\
```{r setup_R, include = F, message = F, warning = F}
# Clear workspace
rm(list=ls())

# Load packages
library(caret)
library(GA)
library(pander)
```

```{r setup_knitr, include = F}
# Set code width to 60 to contain within PDF margins
knitr::opts_chunk$set(tidy = F, tidy.opts = list(width.cutoff = 60))

# Set all figures to be centered
knitr::opts_chunk$set(fig.align = "center")

# Set echo to off
knitr::opts_chunk$set(echo = F)
```

```{r}
#==============================================================================
# Data Import & Prep
#==============================================================================

# Read data
tms = read.csv("C:/Users/michael.gilbert/Dropbox/R/Two_Months_Salary.csv", 
               header = T)

# Recode integers to numeric
tms$price = as.numeric(tms$price)

# Recode integers to factor
tms$color = as.factor(tms$color)
tms$clarity = as.factor(tms$clarity)

# Set factor variable levels
levels(tms$color) = c("D", "E", "F", "G", "H", "I", "J", "K", "L")
levels(tms$clarity) = c("IF", "VVS1", "VVS2", "VS1", "VS2", "SI1", "SI2", 
                        "I1", "I2")

# Rename factor variable levels (replace spaces)
levels(tms$store) = c("Ashford", "Ausmans", "Blue_Nile", "Chalmers", 
                      "Danford", "Fred_Meyer", "Goodmans", "Kay", 
                      "R_Holland", "Riddles", "University", "Zales")
levels(tms$cut) = c("Ideal", "Not_Ideal")

#--------------------------------------
# GitHub
#--------------------------------------
# Create function to source functions from GitHub
source.GitHub = function(url){
    require(RCurl)
    sapply(url, function(x){
        eval(parse(text = getURL(x, followlocation = T,
                                 cainfo = system.file("CurlSSL", 
                                          "cacert.pem", package = "RCurl"))),
             envir = .GlobalEnv)
    })
}

# Assign URL and source functions
url = "http://bit.ly/1T6LhBJ"
source.GitHub(url); rm(url)

#--------------------------------------
# RMSE
#--------------------------------------
rmse = function(pred, actual){
    sqrt(mean((pred - actual)^2))
}

#==============================================================================
# Model Build
#==============================================================================

# Random sample into 70/30 training-test split
set.seed(123)
tms.train = createDataPartition(tms$price, p = 0.70, list = F)
tms.test = as.matrix(as.integer(rownames(tms))[-tms.train])

# Create binary indicator variables
tms = fac.flag(tms)

# Remove factor variables
tms = subset(tms, select = -c(color, clarity, cut, channel, store))

#------------------------------------------------------------------------------
# GA in {GA}
#------------------------------------------------------------------------------

# Build model
tms.m1 = lm(price ~ .,
            data = tms[tms.train, ])
summary(tms.m1)

# Model matrix
x = model.matrix(tms.m1)[, -1]
y = model.response(model.frame(tms.m1))

# Define fitness function (BIC)
fitness = function(string){
    inc = which(string == 1)
    X = cbind(1, x[, inc])
    mod = lm.fit(X, y)
    class(mod) = "lm"
    -BIC(mod)
}

# Create GA model
set.seed(123)
GA = ga("binary",
        fitness = fitness,
        nBits = ncol(x),
        names = colnames(x),
        maxiter = 200)
plot(GA)
summary(GA)

# Build model
tms.m2 = lm(price ~ .,
            data = data.frame(price = y, x[, GA@solution[1, ] == 1]))
summary(tms.m2)

# Build model
tms.m3 = lm(price ~ carat + color_G + color_H + color_F + color_D + color_E + 
                color_I + clarity_SI1 + clarity_VS2 + clarity_SI2 + 
                clarity_I1 + clarity_VS1 + clarity_I2 + cut_Not_Ideal + 
                channel_Mall + store_Goodmans + store_Fred_Meyer,
            data = tms[tms.train, ])
summary(tms.m3)

# In-sample
tms.m3.trn.pred = predict(tms.m3, newdata = tms[tms.train, -2])
tms.m3.trn.rmse = rmse(tms.m3.trn.pred, tms$price[tms.train])

# Out-of-sample
tms.m3.tst.pred = predict(tms.m3, newdata = tms[tms.test, -2])
tms.m3.tst.rmse = rmse(tms.m3.tst.pred, tms$price[tms.test])

#------------------------------------------------------------------------------
# GA in {caret}
#------------------------------------------------------------------------------

# Set fit control
tms.m4.fc = gafsControl(functions = rfGA,
                        method = "cv",
                        number = 5)

# Build model
set.seed(123)
tms.m4 = gafs(x = tms[tms.train, -2],
              y = tms[tms.train, 2],
              iters = 10,
              gafsControl = tms.m4.fc)

# Model summary
tms.m4
tms.m4$ga$final
plot(tms.m4)

# In-sample
tms.m4.trn.pred = predict(tms.m4, newdata = tms[tms.train, -2])
tms.m4.trn.rmse = rmse(tms.m4.trn.pred, tms$price[tms.train])

# Out-of-sample
tms.m4.tst.pred = predict(tms.m4, newdata = tms[tms.test, -2])
tms.m4.tst.rmse = rmse(tms.m4.tst.pred, tms$price[tms.test])

#------------------------------------------------------------------------------
# Results
#------------------------------------------------------------------------------

# Model Names
model.comp = data.frame(Model_Names = c("M3", "M4"),
                        Package = c("{GA}", "{caret}"),
                        Features = c("17", "27"),
                        RMSE_Train = c(tms.m3.trn.rmse, tms.m4.trn.rmse),
                        RMSE_Test = c(tms.m3.tst.rmse, tms.m4.tst.rmse))

pander(model.comp, caption = "", justify = "left")
