---
title: "454_Summary_Function"
author: "Michael Gilbert"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document: default
  html_document:
    fig_caption: yes
    fig_crop: no
    fig_height: 4.75
    fig_width: 5.75
    highlight: tango
geometry: margin = 0.5in
---
\
```{r setup, include = F}
knitr::opts_chunk$set(echo = T)
```

# A Customized Summary Function

Write a R function named my.summary() that will take a data frame of numeric variables and compute summary statistics for the variables. The summary statistics should be the quantiles (0.01,0.05,0.25,0.5,0.75,0.95,0.99), the mean, the variance, the min, and the max. The summary statistics should be output in a R data frame structure. 

To make your function a little more advanced try to include the percent missing.

Apply your function to the wine data set used in Assignment #1. Post an example function call, your function output, and your function.

```{r Summary}
#==============================================================================
# Data Import
#==============================================================================
# Read data
wine = read.csv("~/wine.data", header = F)

# Assign column names
colnames(wine) = c("Class", "Alcohol", "Malic_Acid", "Ash", "Ash_Alcalinity", 
                   "Magnesium", "Phenols_Total", "Flavanoids", 
                   "Phenols_Nonflavanoid", "Proanthocyanins", 
                   "Color_Intensity", "Hue", "OD280_OD315", "Proline")

# Assign wine$class as factor
wine$Class = as.factor(wine$Class)

# Convert integers to numeric
wine$Magnesium = as.numeric(wine$Magnesium)
wine$Proline = as.numeric(wine$Proline)

#==============================================================================
# Function
#==============================================================================
my.summary = function(df, na.rm = T){
    table.summary = data.frame()
    cn = colnames(df[, !sapply(df, is.factor)])
    for (num.var in cn){
        qt = quantile(df[, num.var], na.rm = na.rm, 
                      probs = c(0.01, 0.05, 0.25, 0.50, 0.75, 0.95, 0.99))
        mu = mean(df[, num.var], na.rm = na.rm)
        vr = var(df[, num.var], na.rm = na.rm)
        mn = min(df[, num.var], na.rm = na.rm)
        mx = max(df[, num.var], na.rm = na.rm)
        pm = mean(is.na(df[, num.var])) * 100
        table.row = as.data.frame(cbind(num.var,
                                  round(cbind(t(qt), mu, vr, mn, mx, pm), 
                                  digits = 2)))
        table.summary = rbind(table.summary, table.row)
    }
    colnames(table.summary)[1] = "Variable"
    colnames(table.summary)[9] = "Mean"
    colnames(table.summary)[10] = "Variance"
    colnames(table.summary)[11] = "Minimum"
    colnames(table.summary)[12] = "Maximum"
    colnames(table.summary)[13] = "Percent Missing"
    return(table.summary)
}

#==============================================================================
# Function Call
#==============================================================================
my.summary(wine)
```

```{r FIN}
# Session Info
sessionInfo()
```

