---
title: "454_Change_Machine"
author: "Michael Gilbert"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document: default
  html_document:
    fig_caption: yes
    fig_crop: no
    fig_height: 4.75
    fig_width: 5.75
    highlight: tango
geometry: margin = 0.5in
---
\
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Quick R Knowledge Check

## The Change Machine

Write a R function named change.machine() that takes an input value between 0 and 100 and returns a R list object of the number of quarters, dimes, nickels, and pennies required to provide the correct change. If the user inputs a number not in (0,100), then the function should print an error message to the terminal: 'The input value must be between 0 and 100.' 

```{r Change_Machine}
#==============================================================================
# Function
#==============================================================================
change.machine <- function(val){
    validRange = seq(0, 100)
    if (!val %in% validRange){
        stop("The inpute value must be between 0 and 100.")
    }
    q <- val %/% 25; q.rem <- val %% 25
    d <- q.rem %/% 10; d.rem <- q.rem %% 10
    n <- d.rem %/% 5; n.rem <- d.rem %% 5
    p <- n.rem %/% 1
    change <- list(Quarters = q, Dimes = d, Nickels = n, Pennies = p)
    assign("change", change, envir = .GlobalEnv)
}

#==============================================================================
# Output
#==============================================================================
change.machine(88)
change
```

```{r FIN}
# Session Info
sessionInfo()
```

