---
title: "454_Sample_Flat_File"
author: "Michael Gilbert"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    fig_caption: yes
    fig_crop: no
    fig_height: 4.75
    fig_width: 5.75
    highlight: tango
  pdf_document: default
geometry: margin = 0.5in
---
\
```{r setup, include = F}
knitr::opts_chunk$set(echo = T)
knitr::opts_chunk$set(warning = F) 
knitr::opts_chunk$set(results = "hold")
```

# Sampling a Text File

In the Week #1 module we have a set of practice data manipulation exercises to use as a warm up for the quarter. Most of these exercises are trivial, however #10 is a very important skill. Frequently we will interact with data stored as large text files. As an `R` user we know that the `R` functions `read.table()` and `read.csv()` will read the entire data set into memory, what do we do if we want to access some of the data without reading all of it into memory?

If we wanted to read part of the data in without reading all of the data in, then we would be performing an operation called `chunking`. However, what if we want to read in a random sample. Data is frequently stored in non-random order, and hence we always want to be able to sample a text file 'on the fly' without reading the whole file into memory. How do we do this?

Complete Exercise #10 and post your code on Sunday night so that everyone has a chance to work on the problem.

# Solution

```{r SFF_V1}
#==============================================================================
# Version 1
#==============================================================================
# Set file name
df = "AmesHousing.txt"

# Draw random sample of observation IDs without replacement for specified size 
# Note: row.names = 1 only reads a single column
set.seed(123)
samp = sample(x = nrow(read.table(df, header = T, sep = "\t", row.names = 1)), 
              size = 200, replace = F)

# Extract and concatenate tuples of observation IDs from file
# Note: nrows = 1 only reads a single row
temp = NULL
for (i in samp){
    temp = rbind(temp, read.table(df, header = F, sep = "\t", nrows = 1, 
                 skip = i))
}

# Assign column names
# Note: nrows = 1 only reads a single row
colnames(temp) = colnames(read.table(df, header = T, sep = "\t", nrows = 1))

# Validate sampling did not select duplicate rows
anyDuplicated(temp$SID)
```

```{r SFF_V2, include = F}
#==============================================================================
# Version 2
#==============================================================================
# Set file name
df = "AmesHousing.txt"

# Count rows
rows.count = function(df, head, sep){
    df.rows = 1
    cols = ncol(read.table(df, header = head, sep = sep, nrows = 1))
    repeat{
        tryCatch(read.table(df, header = head, sep = sep, nrows = df.rows, 
                            row.names = 1), 
                 error = function(c){
                     stop(c)
                     })
        df.rows = df.rows + 1
    }
    return(df.rows)
}


# Count rows
rows.count = function(df, head, sep){
    df.rows = 1
    cols = ncol(read.table(df, header = head, sep = sep, nrows = 1))
    repeat{
        temp = tryCatch(read.table(df, header = head, sep = sep, 
                        nrows = df.rows, colClasses = c(rep("integer", 1), 
                        rep("NULL", (cols-1)))), 
                        error = function(c){stop(c)})
        df.rows = df.rows + 1
    }
    return(nrow(temp))
}

# Deploy
rows.count(df, head = T, sep = "\t")


df = "AmesHousing.txt"
df.rows = 2
cols = ncol(read.table(df, header = T, sep = "\t", nrows = 1))
temp = read.table(df, header = T, sep = "\t", nrows = df.rows,
                  colClasses = c(1, rep("NULL", (cols-1))))
nrow(temp)

```

```{r FIN}
# Session Info
sessionInfo()
```
