---
title: "454_Magic_8_Ball"
author: "Michael Gilbert"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    fig_caption: yes
    fig_crop: no
    fig_height: 4.75
    fig_width: 5.75
    highlight: tango
  pdf_document: default
geometry: margin = 0.5in
---
\
```{r setup, include = F}
knitr::opts_chunk$set(echo = T)
```

# Magic 8-ball

Write an `R` function `magic.8ball(` that outputs one the following statements when called.

(1) 'I don't see that happening.'
(2) 'You must be dreaming.'
(3) 'Chances are good.'
(4) 'If you work hard and stay focused, then this might happen.'
(5) 'When the sun shines, it shines on your face.'
(6) 'Definitely going to happen!'
(7) 'Take some time to think about it.'
(8) 'This is not a good idea!'
(9) 'This is a great idea!'
(10) 'Cannot be determined at this time. Try Again!'

__Advanced__ - Include an optional argument named `x.seed` that controls the seed for the random number generator. The default definition of `x.seed` should allow the `magic.8ball()` to be random, but its existence will allow you to manually control the seed value and produce the same output from the `magic.8ball()` each time.

# Function

```{r Magic_8_Ball}
#==============================================================================
# Function
#==============================================================================
magic.8ball = function(x.seed){
    if (!missing(x.seed)){
        set.seed(x.seed)
    }
    sample(c("I don't see that happening.", 
             "You must be dreaming.", 
             "Chances are good.", 
             "If you work hard and stay focused, then this might happen.", 
             "When the sun shines, it shines on your face.", 
             "Definitely going to happen!", 
             "Take some time to think about it.", 
             "This is not a good idea!", 
             "This is a great idea!", 
             "Cannot be determined at this time. Try Again!"), 1)
}
```

# Function Call

```{r FUN_1, results = "hold"}
#==============================================================================
# x.seed value = FALSE
#==============================================================================
magic.8ball()
magic.8ball()
magic.8ball()
```

```{r FUN_2, results = "hold"}
#==============================================================================
# x.seed value = TRUE
#==============================================================================
magic.8ball(x.seed = 123)
magic.8ball(x.seed = 123)
magic.8ball(x.seed = 123)
```

```{r FIN}
# Session Info
sessionInfo()
```
