---
title: "454_Data_Summary"
author: "Michael Gilbert"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document: default
  html_document:
    fig_caption: yes
    fig_crop: no
    fig_height: 4.75
    fig_width: 5.75
    highlight: tango
geometry: margin = 0.5in
---
\
```{r setup, include = F}
knitr::opts_chunk$set(c(echo = T, warning = F, results = "hold"))
```

# Data Summary

This R Programming Challenge is more of a data manipulation exercise than a programming exercise. Here we will use a data file of daily stock prices and volume to exercise some basic needed data skills on the R data frame.

Here are the data exercises:

(1) Read in the data file and check the 'structure' of the data. Is the variable `Date` a factor, string, or date? What data type should it be?

```{r DS.1, indent = "    "}
# Read data
jpm = read.csv("~/JPM.csv", header = T)

# Check date class
class(jpm$Date)
```

    `Date` is read as a `factor` variable by `R`.

(2) Type cast the variable Date into an R date type named TradeDate. Make sure that TradeDate is attached to your data frame.

```{r DS.2, indent = "    "}
jpm$TradeDate = as.Date(jpm$Date, format = "%d-%b-%y")
```

(3) Create a seven character Year-Month string named `YearMonth`. Make sure that `YearMonth` is attached to your data frame.

```{r DS.3, indent = "    "}
jpm$YearMonth = substr(jpm$TradeDate, 1, 7)
```

(4) Compute the total trading volume for each YearMonth.

```{r DS.4, indent = "    "}
round(by(jpm$Volume, jpm$YearMonth, sum), digits = 2)
```

(5) Compute the average closing price for each YearMonth.

```{r DS.5, indent = "    "}
round(by(jpm$Close, jpm$YearMonth, mean), digits = 2)
```

(6) Compute the average daily intraday price spread for each YearMonth. Note that this intraday price spread is computed by the daily high minus the daily low.

```{r DS.6, indent = "    "}
jpm$IDPS = (jpm$High - jpm$Low)
round(by(jpm$IDPS, jpm$YearMonth, mean), digits = 2)
```

(7) Use {base} `R` graphics to make this plot.

```{r DS.7, indent = "    "}
boxplot(jpm$IDPS ~ jpm$YearMonth, las = 2, 
        main = "Boxplots of Intraday Price Spread")
```

(8) Use {base} `R` graphics to make this plot.

```{r DS.8, indent = "    "}
bp = barplot(by(jpm$Close, jpm$YearMonth, mean), las = 2, 
             main = "Average Closing Price",
             ylim = c(0, 80))
abline(h = mean(jpm$Close), col = "blue", lty = 5)
rm(bp)
```
