---
title: "454_Functions"
author: "Michael Gilbert"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document: default
  html_document:
    fig_caption: yes
    fig_crop: no
    fig_height: 4.75
    fig_width: 5.75
    highlight: tango
geometry: margin = 0.5in
---
\
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Quick R Knowledge Check

## The Change Machine

Write a R function named change.machine() that takes an input value between 0 and 100 and returns a R list object of the number of quarters, dimes, nickels, and pennies required to provide the correct change. If the user inputs a number not in (0,100), then the function should print an error message to the terminal: 'The input value must be between 0 and 100.' 

```{r}
#==============================================================================
# Function
#==============================================================================
change.machine <- function(val){
    validRange = seq(0, 100)
    if (!val %in% validRange){
        stop("The inpute value must be between 0 and 100.")
    }
    q <- val %/% 25; q.rem <- val %% 25
    d <- q.rem %/% 10; d.rem <- q.rem %% 10
    n <- d.rem %/% 5; n.rem <- d.rem %% 5
    p <- n.rem %/% 1
    assign("change", list(Quarters = q, Dimes = d, Nickels = n, Pennies = p), 
           envir = .GlobalEnv)
    print(change)
}

#==============================================================================
# Output
#==============================================================================
change.machine(88)
```

## A Customized Summary Function

Write a R function named my.summary() that will take a data frame of numeric variables and compute summary statistics for the variables. The summary statistics should be the quantiles (0.01,0.05,0.25,0.5,0.75,0.95,0.99), the mean, the variance, the min, and the max. The summary statistics should be output in a R data frame structure. 

```{r}
#==============================================================================
# Function
#==============================================================================
my.summary <- function(df){
    table.summary <- data.frame()
    cn <- colnames(df[, !sapply(df, is.factor)])
    for (num.var in cn){
        qt <- quantile(df[, num.var],
                       probs = c(0.01, 0.05, 0.25, 0.50, 0.75, 0.95, 0.99))
        mu <- mean(df[, num.var])
        vr <- var(df[, num.var])
        mn <- min(df[, num.var])
        mx <- max(df[, num.var])
        name.var <- paste(deparse(substitute(df)), "$", num.var, sep = "")
        table.row <- as.data.frame(cbind(name.var,
                                         round(cbind(t(qt), mu, vr, mn, mx), 
                                         digits = 2)))
        table.summary <- rbind(table.summary, table.row)
    }
    rownames(table.summary) <- table.summary[, 1]
    table.summary <- subset(table.summary, select = -name.var)
    colnames(table.summary)[8] <- "Mean"
    colnames(table.summary)[9] <- "Variance"
    colnames(table.summary)[10] <- "Minimum"
    colnames(table.summary)[11] <- "Maximum"
    assign("table.summary", table.summary, envir = .GlobalEnv)
}

#==============================================================================
# Output
#==============================================================================
my.summary(mtcars)
table.summary
```

```{r FIN}
# Session Info
sessionInfo()
```
