---
title: "R_EDA_Sandbox_Demo"
author: "Michael Gilbert"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    fig_caption: yes
    fig_crop: no
    fig_height: 5.00
    fig_width: 6.00
    highlight: tango
  pdf_document: default
geometry: margin = 0.5in
---

```{r setup_knitr, include = F}
# Set code width to 60 to contain within PDF margins
knitr::opts_chunk$set(tidy = F, tidy.opts = list(width.cutoff = 60))

# Set all figures to be centered
knitr::opts_chunk$set(fig.align = "center")

# Set warnings to false
knitr::opts_chunk$set(warning = F)

# Set messages to false
knitr::opts_chunk$set(message = F)
```

\ 

This is a demo of functions I wrote to use in intial `EDA`. The function code is included here as an appendix, though the most recent version is centrally stored on `GitHub` [here](https://github.com/jmichaelgilbert/mspaScripts/blob/master/R_EDA_Sandbox_Functions.R). 

```{r setup_FUN, results = "hide"}
# Load libraries
library(RCurl)
library(pander)

#==============================================================================
# Functions
#==============================================================================
# Create function to source functions from GitHub
source.GitHub = function(url){
    require(RCurl)
    sapply(url, function(x){
        eval(parse(text = getURL(x, followlocation = T,
                                 cainfo = system.file("CurlSSL", 
                                          "cacert.pem", package = "RCurl"))),
             envir = .GlobalEnv)
    })
}

# Assign URL and source functions
url = "http://bit.ly/1T6LhBJ"
source.GitHub(url); rm(url)
```

\ 

# Background

This demo uses the `Auto` dataset from the `ISL` website (the version included in the `{ISLR}` package is slightly different). We'll start by importing these data and doing some minor tweaks based on a data quality check. Our response attribute is `mpg`, but we are also interested exploring data by the class `origin`. 

```{r}
# Download and assign data
if(!file.exists("~/Auto.csv")){
    URL = getURL("http://www-bcf.usc.edu/~gareth/ISL/Auto.csv")
    auto = read.csv(textConnection(URL), header = T)
    rm(URL)
}
```

\ 

# Data Quality Check

The function `str()` helps us check the dimensions, the first few observations, and attribute classes: 

```{r}
# View structure
str(auto)
```

Next we check our summary statistics:

```{r}
# View summary statistics
summary(auto)
```

We notice a few potential issues:

* `cylinders` should probably be a `factor` attribute;
* `horsepower` should probably be a `numeric` attribute;
* `weight` should probably be a `numeric` attribute (`R` will automatically convert `integer` class to `numeric` for mathematical operations);
* `origin` should probably be a `factor` attribute as it indicates country of origin;
* `name` should probably be a `character` attribute.

Additionally, `horsepower` has a `?`. That should probably be a `NA` value (in case any other attributes use the `?` character to denote `NA`, we'll apply it to the whole dataset). Note that with `horsepower`, we first convert it to a character to preserve the numeric values. Let's make these changes:

``` {r}
# Treat "?" as NA
auto[auto == "?"] = NA

# Convert attributes as appropriate
auto$horsepower = as.numeric(as.character(auto$horsepower))
auto$weight = as.numeric(auto$weight)
auto$cylinders = as.factor(auto$cylinders)
auto$origin = as.factor(auto$origin)
auto$name = as.character(auto$name)
```

For the purposes of this demo, we'll drop the `name` attribute and remove the tuples with `NA` values (this is revisited in the _Attribute Manipulations_ section). We'll create a version of `auto` called `auto.am`as a reference for use in this later section.

```{r}
# Drop name attribute
auto = subset(auto, select = -name)

# Create separate data.frame for auto
auto.am = auto

# Remove NA values
auto = na.omit(auto)
```

\ 

# Qualitative EDA

We'll start with some visual `EDA`. The first user-defined function we'll use is `num.plots()`. This function produces a `histogram`, `boxplot`, `scatterplot`, and `Q-Q plot` for each `numeric` attribute _or_ an individual attribute named in `num.plots(df = )`. For the `scatterplot`, the response attribute `mpg` must be named, and the `boxplot` can be split by a factor class. The histogram can also plot bins of _equally spaced probabilities_ (`num.plots(prob = T)`), and the normal curve overlay can be turned on or off (`num.plots(norm = F)`).

The function is really composed of four separate functions, which can each be called individually: `num.hist`, `num.boxplot`, `num.scatter`, and `num.qq`.

```{r, fig.height = 7.00, fig.width = 8.00}
# Create plots of numeric attributes
# Use [mpg] as response in scatterplot
# Use [origin] as factor to split in boxplots
num.plots(df = auto, df.num = auto$mpg, df.fac = auto$origin)
```

We'll continue with your visual `EDA`, but now for `factor` attributes. This uses two user-defined functions: `fac.barplot`, and `fac.mosaic`. The former makes barplots of factor attributes (and will also accept a numeric attribute to segregate on). The latter makes mosaic plots for a named factor. Both will also accept an individual attribute named in `num.plots(df = )` - this is done with the `fac.mosaic()` function.

Note the mosaic plot shows some pretty unbalanced classes when it comes to `cylinders` by `origin`.

```{r}
# Create barplots for factor attributes
fac.barplot(df = auto)

# Create mosaic plot for 
fac.mosaic(df = auto$origin, df.fac.cn = auto$cylinders)
```

\ 

# Quantitative EDA

Now we'll switch gears (no pun intended) over to quantitative `EDA`. First, we'll look at summary statistics for `numeric` attributes split by a named `factor` attribute (`origin`). This user-defined function is called `num.freq()`. By default, the function will return summary stats for _all_ `numeric` (or `integer`) attributes. A second argument, `num.freq(df.num.cn = )`, can be supplied specifying a single `numeric` or `integer` attribute to use. Here we use `mpg`.

```{r}
# Summary stats for numeric attributes, split by named factor
num.freq(df.fac = auto$origin, df.num.cn = auto$mpg)
```

Similarly, we can look at occurence rates by `factor` class - think of this as the quantitative version of the mosaic plot. This user-defined function is called `fac.freq()`. A second argument, `num.freq(df.num.cn = )`, can be supplied specifying a single `factor` attribute to use. A third argument, `fac.freq(cat = )`, can be supplied, returning counts and frequencies of the `factor` (the default is `FALSE`).

```{r}
# Frequency of occurrence for factor attributes, split by named factor
fac.freq(df.fac = auto$origin)

# Counts and frequency of occurrence of named factor
fac.freq(df.fac = auto$origin, cat = F)
```

\ 

# Attribute Manipulations

The last part of the demo deals with various attribute manipulations: missing flag creation, scaling, trimming, transforming, and creating indicator attributes for each `factor` attribute level.

## Missing Values

The user-defined function `miss.flag()` creates indicator attributes for attributes with missing data, as sometimes missingness is predictive. The function defaults to automatically do this for all attributes in the dataset, but can be specified to only do it for a specific attribute, or type (e.g. `numeric` or `factor`). 

```{r}
# Create temp data.frame
auto.temp = auto.am

# Apply missing flag function
auto.temp = miss.flag(df = auto.temp)
```

When viewing attribute names, we see there's a new attribute called `MF_horsepower`, with class factor. 

```{r}
# View names
names(auto.temp)

# Check class
class(auto.temp$MF_horsepower)

# Validate application
sum(is.na(auto.temp$horsepower))
length(auto.temp$MF_horsepower[auto.temp$MF_horsepower == 1])
```

```{r, include = F}
auto.am = auto
```

## Scales, Trims, Transforms, and Indicators

Moving more toward data mining, we might want to scale, trim, and transform attributes en masse. The next user-defined functions do just that: `num.scale()`, `num.trims()`, `num.trans()`. 

### num.scale()

`num.scale()` will center and scale all `numeric` attributes in the dataset, _or_ an individual attribute named in `num.scale(df = )`. Note: the final suffix is `.V1` as each attribute has two additional attributes (`scaled:center` and `scaled:scale`). 

```{r}
# Create temp data.frame
auto.temp = auto.am

# Apply scale function
auto.temp = num.scale(df = auto.temp)
```

When viewing attribute names, we see all `numeric` attributes have been centered and scaled, with the suffix `_cs`. Results can be seen with the `summary()` function.

```{r}
# View names
colnames(auto.temp[, 9:ncol(auto.temp)])

# Summary stats
summary(auto.temp[, 9:ncol(auto.temp)])
```

### num.trims()

`num.trims()` will Winsorize all `numeric` attributes in the dataset, _or_ an individual attribute named in `num.trims(df = )`. The attributes are Winsorized at the 1st/99th, 5th/95th, 10th/90th, and 25th/75th percentiles. 

```{r}
# Create temp data.frame
auto.temp = auto.am

# Apply scale function
auto.temp = num.trims(df = auto.temp)
```

When viewing attribute names, we see all `numeric` attributes have been Winsorized, with the suffixes `_T99`, `_T95`, `_T90`, and `_T75`. Results can be seen with the `summary()` function (trims can be seen at `_T75`).

```{r}
# View names
colnames(auto.temp[, 9:ncol(auto.temp)])

# Summary stats
summary(auto.temp[, 9:ncol(auto.temp)])
```

### num.trans()

`num.trans()` will transform all `numeric` attributes in the dataset, _or_ an individual attribute named in `num.trans(df = )` by taking the natural log (a `+1` is added for stability), square root, and square. 

```{r}
# Create temp data.frame
auto.temp = auto.am

# Apply scale function
auto.temp = num.trans(df = auto.temp)
```

When viewing attribute names, we see all `numeric` attributes have been transformed, with the suffixes `_ln` ($\log_{x}$), `_rt` ($\sqrt{x}$), and `_sq` ($X^{2}$). Results can be seen with the `summary()` function.

```{r}
# View names
colnames(auto.temp[, 9:ncol(auto.temp)])

# Summary stats
summary(auto.temp[, 9:ncol(auto.temp)])
```

### fac.flag()

`fac.flag()` will transform all levels of `factor` attributes in the dataset to indicator attributes, _or_ an individual attribute named in `fac.flag(df = )`. Note: original attributes are not removed.

```{r}
# Create temp data.frame
auto.temp = auto.am

# Apply scale function
auto.temp = fac.flag(df = auto.temp)
```

When viewing attribute names, we see all `factor` attributes have been transformed, with the suffixes `_LEVEL`, where `LEVEL` is a named level in the `factor`. Results can be seen with the `summary()` function.

```{r}
# View names
colnames(auto.temp[, 9:ncol(auto.temp)])

# Summary stats
summary(auto.temp[, 9:ncol(auto.temp)])
```

\ 

\pagebreak

# Appendix: Function Code

```{r, eval = F}
###############################################################################
# R_EDA_Sandbox_Functions.R
# Last updated: 2016-08-02 by MJG
###############################################################################

# A compilation of useful functions to [ideally] deploy on any data set

# Suggested order of deployment:
#   Convert variables as necessary (e.g. to factors)
#   Plots for EDA on numeric and factor variables
#   Missing flags
#   Missing imputes
#   Trims
#   Transforms

#==============================================================================
# Accuracy
#==============================================================================

#--------------------------------------
# fit()
#--------------------------------------
# Function to add MSE to other measures from forecast::accuracy
fit = function(f, x){
    require(forecast)
    temp = data.frame(forecast::accuracy(f, x),
                      forecast::accuracy(f, x)[, 2]^2)
    temp = temp[, -c(1)]
    colnames(temp)[6] <- "MSE"
    temp = temp[c(6, 1, 2, 3, 4, 5)]
    print(temp)
}

#==============================================================================
# Missing Observations
#==============================================================================

#--------------------------------------
# miss.flag()
#--------------------------------------
# Function to create indicator variables as missing flags
miss.flag = function(df, df.cn = c("num", "fac")){
    # Check for columns to apply
    if (missing(df.cn)){
        cols = colnames(df)
    } else if (df.cn == "num"){
        cols = colnames(df[, !sapply(df, is.factor)])
    } else if (df.cn == "fac"){
        cols = colnames(df[, sapply(df, is.factor)])
    }
    # Apply function
    for (i in cols){
        if (sum(is.na(df[, i])) > 0){
            df[paste("MF", i, sep = "_")] =
                as.factor(ifelse(is.na(df[, i]), 1, 0))
        }
    }
    return(df)
}

#==============================================================================
# Numeric Variables
#==============================================================================

#------------------------------------------------------------------------------
# Plots
#------------------------------------------------------------------------------

#--------------------------------------
# num.boxplot()
#--------------------------------------
# Function to create boxplots of numeric variables
num.boxplot = function(df, df.fac){
    # Check for data.frame or attribute
    if (grepl("$", deparse(substitute(df)), fixed = T)){
        if (!class(df) %in% c("numeric", "integer")){
            stop("Please supply a numeric or integer variable to df")
        }
        temp = unlist(strsplit(deparse(substitute(df)),
                               split = "$", fixed = T))
        df = eval(as.name(paste(temp[1])))
        df.name = temp[1]
        cols = temp[2]
    } else {
        df.name = deparse(substitute(df))
        cols = colnames(df[, !sapply(df, is.factor)])
    }
    # Create plot(s)
    for (i in cols){
        if (missing(df.fac)){
            boxplot(df[, i], col = "grey", horizontal = T,
                    main = paste("Boxplot of ", df.name, "$", i, sep = ""),
                    xlab = paste(df.name, "$", i, sep = ""),
                    ylab = "Values")
        } else if (!class(df.fac) %in% c("factor")){
            stop("Please supply a factor variable to df.fac")
        } else {
            fac = unlist(strsplit(deparse(substitute(df.fac)),
                                  split = "$", fixed = T))[2]
            boxplot(df[, i] ~ df[, fac], col = "grey", horizontal = T,
                    main = paste(df.name, "$", i," versus ",
                                 deparse(substitute(df.fac)), sep = ""),
                    ylab = "Values")
        }
    }
}

#--------------------------------------
# num.hist()
#--------------------------------------
# Function to create histograms of numeric variables
# Optional choice of equal spaced probabilities or normal curve overlay
num.hist = function(df, prob = F, norm = T){
    # Check for data.frame or attribute
    if (grepl("$", deparse(substitute(df)), fixed = T)){
        if (!class(df) %in% c("numeric", "integer")){
            stop("Please supply a numeric or integer variable to df")
        }
        temp = unlist(strsplit(deparse(substitute(df)),
                               split = "$", fixed = T))
        df = eval(as.name(paste(temp[1])))
        df.name = temp[1]
        cols = temp[2]
    } else {
        df.name = deparse(substitute(df))
        cols = colnames(df[, !sapply(df, is.factor)])
    }
    # Create plot(s)
    for (i in cols){
        main = paste("Histogram of ", df.name, "$", i, sep = "")
        sub = ifelse(norm, "normal curve overlay (blue)", "")
        y = hist(df[, i], plot = F)
        if (prob){
            seq = seq(0.0, 1.0, by = 0.1)
            h = hist(df[, i], col = "grey", main = main, sub = sub,
                     breaks = quantile(df[, i], probs = seq),
                     xlab = paste(df.name, "$", i, sep = ""))
        }
        if (!prob){
            h = hist(df[, i], col = "grey", main = main, sub = sub,
                     ylim = c(0, 1.15*max(y$counts)),
                     xlab = paste(df.name, "$", i, sep = ""))
        }
        if (norm){
            xfit = seq(min(df[, i]), max(df[, i]), length = 100)
            yfit = dnorm(xfit, mean = mean(df[, i]), sd = sd(df[, i]))
            if (norm & !prob){
                yfit = yfit * diff(h$mids[1:2]) * length(df[, i])
            }
            lines(xfit, yfit, col = "blue", lwd = 2)
        }
    }
}

#--------------------------------------
# num.qq()
#--------------------------------------
# Function to create Q-Q plots of numeric variables
num.qq = function(df){
    # Check for data.frame or attribute
    if (grepl("$", deparse(substitute(df)), fixed = T)){
        if (!class(df) %in% c("numeric", "integer")){
            stop("Please supply a numeric or integer variable to df")
        }
        temp = unlist(strsplit(deparse(substitute(df)),
                               split = "$", fixed = T))
        df = eval(as.name(paste(temp[1])))
        df.name = temp[1]
        cols = temp[2]
    } else {
        df.name = deparse(substitute(df))
        cols = colnames(df[, !sapply(df, is.factor)])
    }
    # Create plot(s)
    for (i in cols){
        qqnorm(df[, i], pch = 21, bg = "grey",
               main = paste("Normal Q-Q Plot of ", df.name, "$", i, sep = ""))
        qqline(df[, i], lwd = 2, col = "blue")
    }
}

#--------------------------------------
# num.scatter()
#--------------------------------------
# Function to create scatterplots of numeric variables
num.scatter = function(df, df.num){
    # Check for data.frame or attribute
    if (grepl("$", deparse(substitute(df)), fixed = T)){
        if (!class(df) %in% c("numeric", "integer")){
            stop("Please supply a numeric or integer variable to df")
        }
        temp = unlist(strsplit(deparse(substitute(df)),
                               split = "$", fixed = T))
        df = eval(as.name(paste(temp[1])))
        df.name = temp[1]
        cols = temp[2]
    } else {
        df.name = deparse(substitute(df))
        cols = colnames(df[, !sapply(df, is.factor)])
    }
    # Create plot(s)
    num = unlist(strsplit(deparse(substitute(df.num)),
                          split = "$", fixed = T))[2]
    for (i in cols){
        plot(df[, i], df[, num], pch = 21, bg = "grey",
             main = paste(df.name, "$", num, " versus ",
                          df.name, "$", i, sep = ""),
             ylab = paste(df.name, "$", num, sep = ""),
             xlab = paste(df.name, "$", i, sep = ""))
    }
}

#--------------------------------------
# num.plots()
#--------------------------------------
# Function to produce four plots per variable:
# num.plots(which = ) corresponds as follows:
#   1 = Histogram
#   2 = Scatterplot
#   3 = Boxplot
#   4 = QQ Plot
num.plots = function(df, df.num, df.fac, prob = F, norm = T,
                     which = c(1, 2, 3, 4)){
    # Check for which plots to create
    if (missing(which)){
        par(mfcol = c(2, 2))
    }
    # Check for data.frame or attribute
    if (grepl("$", deparse(substitute(df)), fixed = T)){
        if (!class(df) %in% c("numeric", "integer")){
            stop("Please supply a numeric or integer variable to df")
        }
        temp = unlist(strsplit(deparse(substitute(df)),
                               split = "$", fixed = T))
        df = eval(as.name(paste(temp[1])))
        df.name = temp[1]
        cols = temp[2]
    } else {
        df.name = deparse(substitute(df))
        cols = colnames(df[, !sapply(df, is.factor)])
    }
    # Create plot(s)
    for (i in cols){
        #------------------------------
        # Histograms
        #------------------------------
        if (1 %in% which){
            main = paste("Histogram of ", df.name, "$", i, sep = "")
            sub = ifelse(norm, "normal curve overlay (blue)", "")
            y = hist(df[, i], plot = F)
            if (prob){
                seq = seq(0.0, 1.0, by = 0.1)
                h = hist(df[, i], col = "grey", main = main, sub = sub,
                         breaks = quantile(df[, i], probs = seq),
                         xlab = paste(df.name, "$", i, sep = ""))
            }
            if (!prob){
                h = hist(df[, i], col = "grey", main = main, sub = sub,
                         ylim = c(0, 1.15*max(y$counts)),
                         xlab = paste(df.name, "$", i, sep = ""))
            }
            if (norm){
                xfit = seq(min(df[, i]), max(df[, i]), length = 100)
                yfit = dnorm(xfit, mean = mean(df[, i]), sd = sd(df[, i]))
                if (norm & !prob){
                    yfit = yfit * diff(h$mids[1:2]) * length(df[, i])
                }
                lines(xfit, yfit, col = "blue", lwd = 2)
            }
        }
        #------------------------------
        # Scatterplots
        #------------------------------
        if (2 %in% which){
            num = unlist(strsplit(deparse(substitute(df.num)),
                                  split = "$", fixed = T))[2]
            plot(df[, i], df[, num], pch = 21, bg = "grey",
                 main = paste(df.name, "$", num, " versus ",
                              df.name, "$", i, sep = ""),
                 ylab = paste(df.name, "$", num, sep = ""),
                 xlab = paste(df.name, "$", i, sep = ""))
        }
        #------------------------------
        # Boxplots
        #------------------------------
        if (3 %in% which){
            if (missing(df.fac)){
                boxplot(df[, i], col = "grey", horizontal = T,
                        main = paste("Boxplot of ", df.name, "$", i, sep = ""),
                        xlab = paste(df.name, "$", i, sep = ""),
                        ylab = "Values")
            } else if (!class(df.fac) %in% c("factor")){
                stop("Please supply a factor variable to df.fac")
            } else {
                fac = unlist(strsplit(deparse(substitute(df.fac)),
                                      split = "$", fixed = T))[2]
                boxplot(df[, i] ~ df[, fac], col = "grey", horizontal = T,
                        main = paste(df.name, "$", i," versus ",
                                     deparse(substitute(df.fac)), sep = ""),
                        ylab = "Values")
            }
        }
        #------------------------------
        # QQ Plots
        #------------------------------
        if (4 %in% which){
            qqnorm(df[, i], pch = 21, bg = "grey",
                   main = paste("Normal Q-Q Plot of ", df.name, "$", i, sep = ""))
            qqline(df[, i], lwd = 2, col = "blue")
        }
    }
    return(par(mfcol = c(1, 1)))
}

#------------------------------------------------------------------------------
# Variable Manipulation
#------------------------------------------------------------------------------

#--------------------------------------
# num.freq()
#--------------------------------------
# Summary stats for numeric variables, split by named factor
num.freq = function(df.fac, df.num.cn){
    table.results = data.frame()
    # Check df.fac is factor
    if (!class(df.fac) %in% c("factor")){
        stop("Please supply a factor variable to df.fac")
    }
    # Assign data.frame and name
    temp = unlist(strsplit(deparse(substitute(df.fac)),
                           split = "$", fixed = T))
    df = eval(as.name(paste(temp[1])))
    fac = temp[2]
    if (missing(df.num.cn)){
        cols = colnames(df[, !sapply(df, is.factor)])
    } else if (!class(df.num.cn) %in% c("numeric", "integer")){
        stop("Please supply a numeric or integer variable to df.num.cn")
    } else {
        cols = unlist(strsplit(deparse(substitute(df.num.cn)),
                               split = "$", fixed = T))[2]
    }
    for (i in cols){
        name.var = rep(paste(i), each = nlevels(df[, fac]))
        name.split = rep(paste(fac), each = nlevels(df[, fac]))
        table.level = levels(df[, fac])
        table.agg = format(aggregate(df[, i], by = list(Var = df[, fac]),
                                     summary)$x, nsmall = 2)
        table.row = as.data.frame(cbind(name.var, name.split,
                                        table.level, table.agg))
        table.results = rbind(table.results, table.row)
    }
    colnames(table.results)[1] = "Variable"
    colnames(table.results)[2] = "Split On"
    colnames(table.results)[3] = "Levels"
    return(table.results)
}

#--------------------------------------
# num.scale()
#--------------------------------------
# Function to scale (normalize: mean = 0, sd = 1) numeric variables
num.scale = function(df){
    # Check for data.frame or attribute
    if (grepl("$", deparse(substitute(df)), fixed = T)){
        if (!class(df) %in% c("numeric", "integer")){
            stop("Please supply a numeric or integer variable to df")
        }
        temp = unlist(strsplit(deparse(substitute(df)),
                               split = "$", fixed = T))
        df = eval(as.name(paste(temp[1])))
        cols = temp[2]
    } else {
        cols = colnames(df[, !sapply(df, is.factor)])
    }
    # Apply function
    for (i in cols){
        i_cs = paste(i, "cs", sep = "_")
        df[i_cs] = scale(df[, i])
    }
    return(df)
}

#--------------------------------------
# num.trans()
#--------------------------------------
# Function to transform numeric variables
num.trans = function(df){
    # Check for data.frame or attribute
    if (grepl("$", deparse(substitute(df)), fixed = T)){
        if (!class(df) %in% c("numeric", "integer")){
            stop("Please supply a numeric or integer variable to df")
        }
        temp = unlist(strsplit(deparse(substitute(df)),
                               split = "$", fixed = T))
        df = eval(as.name(paste(temp[1])))
        cols = temp[2]
    } else {
        cols = colnames(df[, !sapply(df, is.factor)])
    }
    # Apply function
    for (i in cols){
        # Natural Log
        i_ln = paste(i, "ln", sep = "_")
        df[i_ln] = (sign(df[, i]) * log(abs(df[, i])+1))
        # Square Root
        i_rt = paste(i, "rt", sep = "_")
        df[i_rt] = (sign(df[, i]) * sqrt(abs(df[, i])+1))
        # Square
        i_sq = paste(i, "sq", sep = "_")
        df[i_sq] = (df[, i] * df[, i])
    }
    return(df)
}

#--------------------------------------
# num.trims()
#--------------------------------------
# Function to trim numeric variables at various percentiles
num.trims = function(df){
    require(scales)
    # Check for data.frame or attribute
    if (grepl("$", deparse(substitute(df)), fixed = T)){
        if (!class(df) %in% c("numeric", "integer")){
            stop("Please supply a numeric or integer variable to df")
        }
        temp = unlist(strsplit(deparse(substitute(df)),
                               split = "$", fixed = T))
        df = eval(as.name(paste(temp[1])))
        cols = temp[2]
    } else {
        cols = colnames(df[, !sapply(df, is.factor)])
    }
    # Apply function
    for (i in cols){
        # 1st and 99th
        T99 = quantile(df[, i], c(0.01, 0.99))
        df[paste(i, "T99", sep = "_")] = squish(df[, i], T99)
        # 5th and 95th
        T95 = quantile(df[, i], c(0.05, 0.95))
        df[paste(i, "T95", sep = "_")] = squish(df[, i], T95)
        # 10th and 90th
        T90 = quantile(df[, i], c(0.10, 0.90))
        df[paste(i, "T90", sep = "_")] = squish(df[, i], T90)
        # 25th and 75th
        T75 = quantile(df[, i], c(0.25, 0.75))
        df[paste(i, "T75", sep = "_")] = squish(df[, i], T75)
    }
    return(df)
}

#==============================================================================
# Factor Variables
#==============================================================================

#------------------------------------------------------------------------------
# Plots
#------------------------------------------------------------------------------

#--------------------------------------
# fac.barplot()
#--------------------------------------
# Function to create barplots of factor variables
fac.barplot = function(df, df.num){
    # Check for data.frame or attribute
    if (grepl("$", deparse(substitute(df)), fixed = T)){
        if (!class(df) %in% c("factor")){
            stop("Please supply a factor variable to df")
        }
        temp = unlist(strsplit(deparse(substitute(df)),
                               split = "$", fixed = T))
        df = eval(as.name(paste(temp[1])))
        df.name = temp[1]
        cols = temp[2]
    } else {
        df.name = deparse(substitute(df))
        cols = colnames(df[, sapply(df, is.factor)])
    }
    # Create plot(s)
    for (i in cols){
        if (missing(df.num)){
            plot(df[, i],
                 main = paste(df.name, "$", i, sep = ""),
                 ylim = c(0, 1.15*max(summary(df[, i]))),
                 ylab = "Frequency")
        } else if (!class(df.num) %in% c("numeric", "integer")){
            stop("Please supply a numeric variable to df.num")
        } else {
            num = unlist(strsplit(deparse(substitute(df.num)),
                                  split = "$", fixed = T))[2]
            barplot(table(df[, num], df[, i]),
                    main = paste(df.name, "$", i, " by ", 
                                 deparse(substitute(df.num)), sep = ""),
                    ylim = c(0, 1.15*max(table(df[, i], df[, num]))),
                    ylab = "Frequency", beside = T)
        }
    }
}

#--------------------------------------
# fac.mosaic()
#--------------------------------------
# Function to create mosaic plots of factor variables
fac.mosaic = function(df.fac, df.fac.cn){
    require(RColorBrewer)
    # Check df.fac is factor
    if (!class(df.fac) %in% c("factor")){
        stop("Please supply a factor variable to df.fac")
    }
    # Assign data.frame and name
    temp = unlist(strsplit(deparse(substitute(df.fac)),
                           split = "$", fixed = T))
    df = eval(as.name(paste(temp[1])))
    df.name = temp[1]
    fac = temp[2]
    # Check if df.fac.cn is missing or named (and class if named)
    if (missing(df.fac.cn)){
        cols = colnames(df[, sapply(df, is.factor)])
    } else if (!class(df.fac.cn) %in% c("factor")){
        stop("Please supply a factor variable to df.fac.cn")
    } else {
        cols = unlist(strsplit(deparse(substitute(df.fac.cn)),
                               split = "$", fixed = T))[2]
    }
    # Create plot(s)
    for (i in cols){
        plot(df[, fac], df[, i],
             col = brewer.pal(nlevels(df[, i]), "Spectral"),
             main = paste(df.name, "$", fac," versus ",
                          df.name, "$", i, sep = ""),
             xlab = paste(df.name, "$", fac, sep = ""),
             ylab = paste(df.name, "$", i, sep = ""))
    }
}

#------------------------------------------------------------------------------
# Variable Manipulation
#------------------------------------------------------------------------------

#--------------------------------------
# fac.freq()
#--------------------------------------
# Frequency of occurrence for factor variables, split by named factor
fac.freq = function(df.fac, df.fac.cn, cat = T){
    table.results = data.frame()
    # Check df.fac is factor
    if (!class(df.fac) %in% c("factor")){
        stop("Please supply a factor variable to df.fac")
    }
    # Assign data.frame and name
    temp = unlist(strsplit(deparse(substitute(df.fac)),
                           split = "$", fixed = T))
    df = eval(as.name(paste(temp[1])))
    fac = temp[2]
    # Check if df.fac.cn is missing or named (and class if named)
    if (missing(df.fac.cn)){
        cols = colnames(df[, sapply(df, is.factor)])
    } else if (!class(df.fac.cn) %in% c("factor")){
        stop("Please supply a factor variable to df.fac.cn")
    } else {
        cols = unlist(strsplit(deparse(substitute(df.fac.cn)),
                               split = "$", fixed = T))[2]
    }
    # Factor splits
    if (cat){
        for (i in cols){
            name.var = rep(paste(i), each = nlevels(df[, fac]))
            name.split = rep(paste(fac), each = nlevels(df[, fac]))
            table.level = levels(df[, fac])
            table.agg = aggregate(df[, i], by = list(Var = df[, fac]),
                                  summary)$x
            table.prop = format(round(prop.table(table.agg, 1) * 100,
                                      digits = 2), nsmall = 2)
            table.results = as.data.frame(cbind(name.var, name.split,
                                                table.level, table.prop))
            colnames(table.results)[1] = "Variable"
            colnames(table.results)[2] = "Split On"
            colnames(table.results)[3] = "Levels"
            if (missing(df.fac.cn)){
                print(table.results)
            } else {
                return(table.results)
            }
        }
    }
    # Factor counts and frequencies
    if (!cat){
        name.var = rep(paste(fac), each = 2)
        name.type = c("Count", "Percent")
        table.agg = t(summary(df[, fac]))
        table.prop = format(round(prop.table(table.agg) * 100,
                                  digits = 2), nsmall = 2)
        table.row = rbind(table.agg, table.prop)
        table.col = cbind(name.var, name.type, table.row)
        table.results = as.data.frame(table.col)
        colnames(table.results)[1] = "Variable"
        colnames(table.results)[2] = "Type"
        return(table.results)
    }
}

#--------------------------------------
# fac.flag()
#--------------------------------------
# Function to create indicator variables from factor variable levels
fac.flag = function(df){
    # Check for data.frame or attribute
    if (grepl("$", deparse(substitute(df)), fixed = T)){
        if (!class(df) %in% c("factor")){
            stop("Please supply a factor variable to df")
        }
        temp = unlist(strsplit(deparse(substitute(df)),
                               split = "$", fixed = T))
        df = eval(as.name(paste(temp[1])))
        cols = temp[2]
    } else {
        cols = colnames(df[, sapply(df, is.factor)])
    }
    # Apply function
    for (i in cols){
        for (level in unique(df[, i])){
            df[paste(i, level, sep = "_")] =
                as.factor(ifelse(df[, i] == level, 1, 0))
        }
    }
    return(df)
}
```

```{r FIN}
#==============================================================================
# FIN
#==============================================================================
sessionInfo()
```
